### 2023-03-15 outputs/paper/paper.qmd
Commented all lines of code on "paper.qmd".

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Sakura Ariga
  - Aliyah Maxine Ramos
  - Annie Yan
thanks: "Code and data are available at: https://github.com/AnnieYan0807/GSS-data-analysis.git."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Set up workspace 
#install.packages("kableExtra")
#install.packages("tidyverse")
#install.packages("lessR")
#install.packages("tidyr")
#install.packages("reshape2")
#install.packages("knitr")
#install.packages("haven")
#install.packages("vtable")

library(kableExtra)
library(tidyverse)
library(lessR)
library(tidyr)
library(reshape2)
library(knitr)
library(haven)
library(vtable)
```

```{r}
#| include: false

# Import cleaned data from the outputs folder, originally imported via the 01-download_data.R file
cleaned_2016gss_data <- read.csv(here::here("outputs/data/cleaned_2016gss_data.csv"))

## Import averaged variable datasets from the outputs folder, originally created in the 02-data_cleaning.R file
mean_age <- read.csv(here::here("outputs/data/mean_age.csv"))
mean_happy <- read.csv(here::here("outputs/data/mean_happy.csv"))
mean_marital <- read.csv(here::here("outputs/data/mean_marital.csv"))
mean_race <- read.csv(here::here("outputs/data/mean_race.csv"))
mean_sex <- read.csv(here::here("outputs/data/mean_sex.csv"))
mean_childs <- read.csv(here::here("outputs/data/mean_childs.csv"))
```

# 1 Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

# 2 Data {#sec-data}

## 2.1 Survey

Saku :) :)

### 2.1.1 Key Features

Saku

### 2.1.2 Methodology

Saku

### 2.1.3 Strengths & Weaknesses

Saku

## 2.2 Questionnaire

Saku

## 2.3 Overview of Data

### 2.3.1 Data Source

Max

### 2.3.2 Variables of Interest

Max

```{r}
#| echo: false

# Table showing the first 10 rows of our cleaned dataset
cleaned_2016gss_data |> 
  select(wwwhr, age, sex, race, 
         degree, childs, marital,
         happy) |> 
  slice(1:8) |> 
  kable(caption='A subset of key features',
        col.names = c("Weekly Internet Hours", "Age", "Sex", "Race", "Highest Degree", 
                      "Number of Children", "Marital Status", "Happiness Level"),
        digits = 1,
        booktabs = FALSE,
        linesep = "") |>
  kable_styling(full_width = T)
```

Table :)

```{r}
#| label: fig-agesidebyside
#| fig-cap: Caption
#| fig-subcap: ["Subcaption"]
#| layout-ncol: 2
#| echo: false
#| warning: false

# Bar graph showing relationship between age and sex
cleaned_2016gss_data |>
  ggplot(mapping = aes(x = age, fill = sex)) + 
  theme(axis.text.x = element_text(angle = 45,hjust=1)) +
  geom_bar(position = "dodge2")

# Bar graph showing relationship between age and race 
cleaned_2016gss_data |>
  ggplot(mapping = aes(x = age, fill = race)) + 
  theme(axis.text.x = element_text(angle = 45,hjust=1)) +
  geom_bar(position = "dodge2")
```

@fig-agesidebyside

```{r}
#| label: fig-childsandmarital
#| fig-cap: Caption
#| echo: false

# Bar graph showing relationship between number of children and marital status
cleaned_2016gss_data |>
  ggplot(mapping = aes(x = childs, fill = marital)) + 
  geom_bar(position = "dodge2")
```

@fig-childsandmarital

```{r}
#| label: fig-happiness
#| fig-cap: Caption
#| echo: false

# Bar graph showing the different levels of happiness and number of respondents for each response
cleaned_2016gss_data |>
  ggplot(mapping = aes(x = happy)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45,hjust=1)) +
  labs(x = "Level of Happiness", y = "Number of Respondents")
```

@fig-happiness

# 3 Results

```{r}
#| label: fig-ageandinternet
#| fig-cap: Average Weekly Total of Internet Use by Age Group
#| echo: false

# Bar graph showing relationship between age and internet use, using tidyverse-style code to pivot and summarise the data 
mean_age |> ggplot(aes(x = variable, y = value)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45,hjust=1), 
        axis.line = element_line(colour = "black"),
        title = element_text(size = 12)) +
  labs(
    x = "Age Group",
    y = "Average Internet Usage Hours"
  )
```

@fig-ageandinternet depicting the relationship between age and average weekly internet usage. The data on internet usage hours for various age groups are averaged to reduce any bias due to selectively chosen survey responses. The curve on the graph is skewed to the right, with its peak occurring in the 21 to 29 age group, which has the highest internet usage of approximately 18 hours per week. Subsequently, internet usage gradually declines with age, with the lowest usage recorded in the 89+ age group at an average of 2 hours per week. This graph indicates that young adults, particularly those aged 21 to 29, spend more time on the internet. As people age, their internet usage tends to decline gradually. The graph suggests a statistically significant correlation between age and internet usage, and the trend is relatively clear.

```{r}
#| label: fig-genderandinternet
#| fig-cap: Average Weekly Total of Internet Use by Gender
#| echo: false

# Bar graph showing relationship between sex and internet use, using tidyverse-style code to pivot and summarise the data 
mean_sex |> ggplot(aes(x = variable, y = value)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45,hjust=1), 
        axis.line = element_line(colour = "black"),
        title = element_text(size = 12)) +
  labs(
    x = "Sex",
    y = "Average Internet Usage Hours"
  )
```

@fig-genderandinternet illustrates the average weekly internet usage of two genders. Similarly to the age and internet usage graph, the data on internet usage hours for different genders are averaged to minimize any volunteering bias. The data shows that there is no substantial difference between male and female internet usage. Female participants have an average of 12.9 hours per week spent on the internet, while male participants have an slightly higher average of 13.3 hours per week spent on the internet. However, the gender selection in this survey is limited, and further discussion regarding this limitation can be found in the ethics and bias section @sec-bias.

```{r}
#| label: fig-raceandinternet
#| fig-cap: Average Weekly Total of Internet Use by Race
#| echo: false

# Bar graph showing relationship between race and internet use, using tidyverse-style code to pivot and summarise the data 
mean_race |> ggplot(aes(x = variable, y = value)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45,hjust=1), 
        axis.line = element_line(colour = "black"),
        title = element_text(size = 12)) +
  labs(
    x = "Race",
    y = "Average Internet Usage Hours"
  )
```

@fig-raceandinternet displays the average weekly internet usage of individuals belonging to different races. Based on the data, we can observe that individuals who do not identify as either Black or White have the highest internet usage, with an average of 15.5 hours per week. African American individuals have a similar level of internet usage, also averaging 15.2 hours per week. In contrast, white individuals tend to have the lowest internet usage, with an average of 12.3 hours per week.

```{r}
#| label: fig-childrenandinternet
#| fig-cap: Average Weekly Total of Internet Use by Children
#| echo: false

# Bar graph showing relationship between number of children and internet use, using tidyverse-style code to pivot and summarise the data 
mean_childs |> ggplot(aes(x = variable, y = value)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45,hjust=1), 
        axis.line = element_line(colour = "black"),
        title = element_text(size = 12)) +
  labs(
    x = "Number of Children",
    y = "Internet Usage Hours"
  )
```

Max

```{r}
#| label: fig-maritalandinternet
#| fig-cap: Average Weekly Total of Internet Use by Marital Status
#| echo: false

# Bar graph showing relationship between marital status and internet use, using tidyverse-style code to pivot and summarise the data 
mean_marital |> ggplot(aes(x = variable, y = value)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45,hjust=1), 
        axis.line = element_line(colour = "black"),
        title = element_text(size = 12)) +
  labs(
    x = "Marital Status",
    y = "Average Internet Usage Hours"
  )
```

@fig-maritalandinternet displays the relationship between marital status and internet usage in 2016. In accordance with popular opinion, people who have never been married use the Internet the most. Most notably, people who are widowed use the Internet the least, perhaps due to the correlation between age and widowhood.

```{r}
#| label: fig-happyandinternet
#| fig-cap: Average Weekly Total of Internet Use by Level of Happiness
#| echo: false

# Bar graph showing relationship between level of happiness and internet use, using tidyverse-style code to pivot and summarise the data 
mean_happy |> ggplot(aes(x = variable, y = value)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45,hjust=1), 
        axis.line = element_line(colour = "black"),
        title = element_text(size = 12)) +
  labs(
    x = "Level of Happiness",
    y = "Average Internet Usage Hours"
  )
```

@fig-happyandinternet displays the relationship between perceived level of happiness and internet usage in 2016. There seems to be a direct decrease in internet usage the more happy the respondent measures themselves to be, with a difference of 2.5 hours between the happiest respondents and least happy respondents.

# 4 Discussion {#sec-discussion}

## 4.1 First discussion point {#sec-first-point}

## 4.2 Second discussion point {#sec-second-point}

## 4.3 Third discussion point {#sec-third-point}

## 4.4 Weaknesses and next steps {#sec-weaknesses}

## 4.5 Ethics and Bias {#sec-bias}

Based on the data collection methodology provided by GSS, we have identified some potential biases that could arise. The survey's target population consists of adults aged 18 and above who reside in households in the United States. Since this sampling approach focuses only on a specific age group, our analysis may be biased with respect to the age factor. Gss also claimed to be strictly voluntary. While this approach supports research ethics by respecting individuals' autonomy and right to privacy, it can also result in volunteer bias. GSS sample are from a mix of urban, suburban, and rural geographic areas, but only a few thousand respondents are interviewed in the main study. The small sample size may be lead to chance findings. It is likely not statistically significant enough. We also suspect potential sampling bias may screw our end result. However, with not enough information disclosed, we are unable to confirm that. The gender and race options provided by the GSS survey are limited. Participants are only given the options of selecting either "Female" or "Male" for gender, and "Blacks," "Whites," or "Others" for race. This narrow selection does not provide an adequate range of options for survey participants.

# 5 Appendix {.unnumbered}

# 6 Additional details

# 7 References